Numerical trajectory patterns of floating macroalgae

### 2021-01-12 thesis_chapter_3.Rmd
draft version ready

The region of interest for this study was the Cape Peninsula in the Western Cape (Figure 1A). A release site was chosen for this study, Kommetjie, which is located on the Atlantic side of the Cape Peninsula (Figure 2B). The hydrodynamics in the study region are driven by a complex interaction of wind and waves. which are also influenced by larger oceanic processes In South Africa, the larger processes can largely be attributed to the Agulhas Current (AC) and the Benguela Current (BC). The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly south-westward following the continental shelf edge and eventually retroflecting eastward back into the South Indian Ocean.

@@

```{r morphology table, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.cap="Table summarising the atrributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation."}
area_table <- read.csv("tables/chapter_3/area_table.csv", sep = ";")
area_table <- flextable(data = area_table)
area_table <- theme_zebra(area_table, odd_header = "transparent", even_header = "transparent")
area_table <- align(x= area_table, align = "center", part = "all")
area_table <- set_header_labels(area_table, Plant.characteristic = "Plant characteristic", Approximate.shape = "Approximate shape", Equation.used = "Equation Used", Plant.dimensions = "Plant dimensions")
area_table <- set_header_labels(area_table, Ac.Minimum = "Ac Minimum", Ac.Mean = "Ac Mean", Ac.Maximum = "Ac Maximum")
area_table <- set_table_properties(area_table, width = 1, layout = "autofit")
area_table <- set_caption(sim_table, "Table 2:Summary of the atrributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation.")
area_table
```

@@

The treatment of drifting macroalgae as purely lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddies or vortices [@fossette2012; @VanSebille2018; @miron2020]. Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics [@putman2018; @brooks2018;@brooks2019; @putman2020], current research has focused on large macroalgae-rafts which can vary greatly in size and surface area. None of the past research has taken advantage of the advancement in numerical ocean models and lagrangian trajectory modeling to investigate how aspects of drag (hydrodynamic and wind) affect the trajectory patterns of solitary drifting macroalgae. This study compared particles trajectories with simulations that included no forms of drag with simulations that included various forms of both hydrodynamic and wind drag. The mean virtual kelp trajectory deviated from the mean passive  trajectory when including varying levels of hydrodynamic and wind drag scenarios. The results from this study show the inclusion of hydrodynamic and wind drag improves entrainment patterns, and also our understanding of how surface area plays a role in solitary drifting macroalgae trajectory, in this case *E.maxima*.

>>>>

The region of interest for this study was the Cape Peninsula in the Western Cape (Figure 1A). A release site was chosen for this study, Kommetjie, which is located on the Atlantic side of the Cape Peninsula (Figure 2B). The hydrodynamics in the study region are driven by a complex interaction of wind and waves. which are also influenced by larger oceanic processes In South Africa, the larger processes can largely be attributed to the Agulhas Current (AC) and the Benguela Current (BC). The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly south-westward following the continental shelf edge and eventually retroflecting eastward back into the South Indian Ocean. The region where the retroflection occurs is known as the Agulhas retroflection and consists of some of the highest levels of mesoscale variability [@garzoli1996]. The variability results in shedding of rings and filaments known as the Agulhas Rings and have an average diameter of ~324km [@lutjeharms1988; @lutjeharms2006; @@lutjeharms2007]. The Agulhas Rings are the main contributors to the Agulhas leakage which is the transport of warm water from the Indian Ocean to the South Atlantic Ocean [@garzoli1996; @lutjeharms2006; @beal2015]. The Agulhas leakage is also a source of meso-scale variability within the Benguela region as anticyclonic rings are shed and drift northwestward along the southern Benguela slope region [@rubio2009]. The Benguela region is characterised by strong coastal upwelling cells which a primarily driven by equatorward winds [@shannon1996; @hardman2003], and is known as the Benguela upwelling region. The Benguela upwelling region starts at 27$^\circ$S and extends southward to 35$^\circ$S [@shannon1996]. The region is also characterised by high amounts of mesoscale variability in the form of eddies and filaments[@blanke2002; @blanke2005; @rubio2009]. The Benguela inshore region is dominated by a northwestward flow as a result of topographical steering, wind stress and interactions with passing Agulhas rings and eddies [@veitch2010].

@@

```{r morphology table, message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.cap="Table summarising the atrributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation."}
area_table <- read.csv("tables/chapter_3/area_table.csv", sep = ";")
area_table <- flextable(data = area_table)
area_table <- theme_zebra(area_table, odd_header = "transparent", even_header = "transparent")
area_table <- align(x= area_table, align = "center", part = "all")
area_table <- set_header_labels(area_table, Plant.characteristic = "Plant characteristic", Approximate.shape = "Approximate shape", Equation.used = "Equation Used", Plant.dimensions = "Plant dimensions")
area_table <- set_header_labels(area_table, Ac.Minimum = "Ac Minimum", Ac.Mean = "Ac Mean", Ac.Maximum = "Ac Maximum")
area_table <- set_table_properties(area_table, width = 1, layout = "autofit")
area_table <- set_caption(area_table, "Table 2:Summary of the atrributes and estimates used to calculate the overall surface area of Kommetjie kelp individuals used in the various simulation.")
area_table
```

@@

The treatment of drifting macroalgae as purely lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddies or vortices [@fossette2012; @miron2020]. Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics [@putman2018; @brooks2018;@brooks2019; @putman2020], current research has focused on large macroalgae-rafts which can vary greatly in size and surface area. None of the past research has taken advantage of the advancement in numerical ocean models and lagrangian trajectory modeling to investigate how aspects of drag (hydrodynamic and wind) affect the trajectory patterns of solitary drifting macroalgae. This study compared particles trajectories with simulations that included no forms of drag with simulations that included various forms of both hydrodynamic and wind drag. The mean virtual kelp trajectory deviated from the mean passive  trajectory when including varying levels of hydrodynamic and wind drag scenarios. The results from this study show the inclusion of hydrodynamic and wind drag improves entrainment patterns, and also our understanding of how surface area plays a role in solitary drifting macroalgae trajectory, in this case *E.maxima*.
